\chapter*{Acknowledgement}


First and foremost, I want to express my deepest love to my wife, Noushin Parizi. You have been my rock, my confidant, and my greatest source of strength throughout this entire journey. Your endless encouragement, motivation, and love have carried me through the toughest times. Your patience during my illness was a lamppost that guided me through the darkest times. Noushin, without you, this achievement would not have been possible. Your belief in me gave me the strength to keep going, and for that, I am forever grateful. This accomplishment is as much yours as it is mine.

I would like to express my deepest gratitude to my supervisor, Prof. dr. Paul Havinga. Paul was not only a guiding light throughout my research journey but also a mentor and friend. He challenged me to think critically and believed in my potential even when I doubted myself and during my illness. His legacy will continue to live on through the many students he has mentored and the profound impact he has had on the field of pervasive computing. I hope this thesis stands as a testament to his mentorship and our shared dedication to advancing knowledge. Paul, you will be greatly missed, but your influence will forever be etched in my mind and this body of work. Thank you for everything.

I would also like to acknowledge my co-supervisor Dr. Berend Jan van der Zwaag, and Dr. Filipe Braganca for their invaluable assistance. Their knowledge and experience helped me gain a better understanding of the equine biomechanics field. Their insights and guidance helped me lay the foundation for my research.

My sincere thanks go to Jeanne Parmentier, Dr. Ineke Smit, and Dr. Carolien Munsters for their support in data collection and equine-related knowledge. Their expertise and assistance were crucial in achieving the research goals.

I would like to thank Nicole Baveld, Rob Bemthuis, Sabari Anbalagan, Jeroen Klein Brinke, Nikita Sharma, Andreas Kamilaris, and yet again, Jeanne Parmentier for their mental support and for being a constant source of inspiration and motivation throughout this journey.

I am also grateful to all the individuals who provided support and guidance throughout my PhD journey. Your contributions have been invaluable in helping me achieve this milestone in my academic career.